---
title: "Surface_water_data_prep"
author: "Wendel Raymond"
date: "April 1, 2021"
output: html_document
editor_options: 
  chunk_output_type: console
---

## Data preparation of SJC surface water
This script will manipulate San Juan County surface water raw data to tall/tidy format. This will ease archival and analysis.

```{r libraries, echo = FALSE, message = FALSE, warning = FALSE}
library(tidyverse)

theme_set(theme_classic())

```

### Data
```{r}
dat <- read.csv("../Historic data/Surface_Water_2019-2020_ALL 01-20-21_WR.csv", header = TRUE, stringsAsFactors = FALSE)
```

### Make tall data
Levels of "Analyte.Name" need to become column names with "Result" and the filling value. Otherwise the rest of the data are in unique to a sampling instance.

```{r}
dat.tl <- spread(dat[, c(1:8)], Analyte.Name, Result)
```

### Clean up
Fix column names, dates, append other useful data

Date
```{r}
dat.tl$Date.Sampled <- as.Date(dat.tl$Date.Sampled,  format = "%m/%d/%Y")
```

Column names
```{r}
nams <- c("sample_no", "sample_desc", "matrix", "date_sampled", "qualifier", "dilution_factor", "E_coli_MF_CFU100mL", "F.coli_CFU100mL", "total_KN_mgL", "total_N03/N02_mgL", "total_N03+N02_mgL", "total_N_mg/L", "total_P_mgL")

colnames(dat.tl) <- nams
```

### Export
```{r}
write.csv(dat.tl, "../Historic data/SJC_Surface_Water_2019-2020_All.csv", row.names = FALSE)
```

